Antigenic phenotypes and phylogenetic metrics improve forecasts of seasonal influenza A/H3N2 evolution

### 2020-05-22 manuscript/main.tex
Add bootstrap analysis to text and tables
Updates Tables 1 and 2 and Supplemental Table S3 and their captions to include
indicators for models that were significantly closer to the future than the
corresponding populations' naive model. Adds a description of the bootstrap
hypothesis tests to the methods and adds figures displaying actual p-values and
corresponding bootstraps of different models to the supplementary material.

\begin{table}[htb]
  \begin{center}
    \scalebox{0.9}{
      \input{tables/simulated_model_selection.tex}
    }
    \caption{
      Model coefficients and performance on validation and test data for simulated populations ordered from best to worst by distance to the future in the validation analysis.
      Coefficients are the mean $\pm$ standard deviation for each metric in a given model across 33 training windows.
      Distance to the future (mean $\pm$ standard deviation) measures the distance in amino acids between estimated and observed future populations.
      The number of times (and percentage of total times) each model outperformed the naive model measures the benefit of each model over a model than estimates no change between current and future populations.
      Test results are based on 18 timepoints not observed during model training and validation.
    }
    \label{table_simulated_model_selection}
  \end{center}
\end{table}

@@

\begin{table}[htb]
  \begin{center}
    \scalebox{0.9}{
      \input{tables/natural_model_selection.tex}
    }
    \caption{
      Model coefficients and performance on validation and test data for natural populations ordered from best to worst by distance to the future in the validation analysis, as in Table~\ref{table_simulated_model_selection}.
      Validation results are based on 23 timepoints.
      Test results are based on eight timepoints not observed during model training and validation.
    }
    \label{table_natural_model_selection}
  \end{center}
\end{table}

@@

>>>>

\begin{table}[htb]
  \begin{center}
    \scalebox{0.9}{
      \input{tables/simulated_model_selection.tex}
    }
    \caption{
      Model coefficients and performance on validation and test data for simulated populations ordered from best to worst by distance to the future in the validation analysis.
      Coefficients are the mean $\pm$ standard deviation for each metric in a given model across 33 training windows.
      Distance to the future (mean $\pm$ standard deviation) measures the distance in amino acids between estimated and observed future populations.
      Distances annotated with asterisks (*) were significantly closer to the future than the naive model as measured by bootstrap hypothesis tests (see Methods).
      The number of times (and percentage of total times) each model outperformed the naive model measures the benefit of each model over a model than estimates no change between current and future populations.
      Test results are based on 18 timepoints not observed during model training and validation.
    }
    \label{table_simulated_model_selection}
  \end{center}
\end{table}

@@

\begin{table}[htb]
  \begin{center}
    \scalebox{0.9}{
      \input{tables/natural_model_selection.tex}
    }
    \caption{
      Model coefficients and performance on validation and test data for natural populations ordered from best to worst by distance to the future in the validation analysis, as in Table~\ref{table_simulated_model_selection}.
      Distances annotated with asterisks (*) were significantly closer to the future than the naive model as measured by bootstrap hypothesis tests (see Methods).
      Validation results are based on 23 timepoints.
      Test results are based on eight timepoints not observed during model training and validation.
    }
    \label{table_natural_model_selection}
  \end{center}
\end{table}

@@

\subsubsection*{Model comparison by bootstrap hypothesis testing}

We compared the performance of each biologically-informed model to the corresponding population's naive model using bootstrap hypothesis tests \cite{EfronBradley1998}.
For each model's empirical distribution of earth mover's distances to the future, $\mathbf{y}$, and the corresponding naive model's distribution, $\mathbf{x}$, we applied the following procedure to test whether their means were significantly different.
We calculated the test statistic $t$ as follows where $n$ and $m$ were the respective lengths of the naive and biologically-informed model's distributions.

$$
t = \frac{\bar{x}-\bar{y}}{\sqrt{\sigma_x^2/n + \sigma_y^2/m}}
$$

Next, we combined the two empirical distributions of distances into a single distribution, $\mathbf{z}$, and calculated the mean of this distribution.
For each original distribution, we created a new distribution that was recentered on the mean of $\mathbf{z}$.
For example, we recentered the $\mathbf{x}$ distribution as $\mathbf{x^{'}} = \mathbf{x} - \bar{x} + \bar{z}$.
We drew $n$ and $m$ samples, respectively, from the resulting distributions, $\mathbf{x^{'}}$ and $\mathbf{y^{'}}$, with replacement for 10,000 bootstrap iterations.
For each iteration, we recalculated the test statistic $t^{'}$ for the means and variances of the bootstrap samples.
We calculated the p-value as the proportion of bootstrap samples for which the $t^{'}$ statistic was greater than or equal to the observed $t$ for the empirical distributions.
To assist with the interpretation of these p-values, we further bootstrapped each model's empirical distance distribution 10,000 times, calculated the mean of each bootstrap sample, and visualized the distributions of each model annotated by its p-value from the test above (Supplemental Figs.~\ref{sup_fig:bootstrap_distributions_for_simulated_sample_3} and \ref{sup_fig:bootstrap_distributions_for_natural_sample_1_with_90_vpm_sliding}).

### 2020-05-22 manuscript/supplement.tex
Add bootstrap analysis to text and tables
Updates Tables 1 and 2 and Supplemental Table S3 and their captions to include
indicators for models that were significantly closer to the future than the
corresponding populations' naive model. Adds a description of the bootstrap
hypothesis tests to the methods and adds figures displaying actual p-values and
corresponding bootstraps of different models to the supplementary material.

@@

@@

\begin{table}[H]
  \begin{center}
    \scalebox{0.9}{
        \input{tables/complete_natural_model_selection.tex}
    }
    \caption{
      All model coefficients and performance on validation and test data for natural populations ordered from best to worst by distance to the future, as in Table~\ref{table_simulated_model_selection}.
      Validation results are based on 23 timepoints.
      Test results are based on eight timepoints not observed during model training and validation.
      Model results for additional variants of fitness metrics including those based on epitope mutations and DMS preferences are included for reference.
    }
    \label{sup_table:complete_natural_model_selection}
  \end{center}
\end{table}

>>>>

\begin{figure}[H]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/bootstrap_distributions_for_simulated_sample_3.pdf}
  \caption{
  Bootstrap distributions of the mean distance to the future for simulated populations.
  Empirical distances to the future were sampled with replacement and mean values for each bootstrap sample were calculated across n=10,000 bootstrap iterations.
  The horizontal gray line indicates the median of the bootstrap distribution for the naive model.
  Each model is annotated by the p-value from a bootstrap hypothesis test between the mean distances for that model and the naive model (see Methods).
  }
  \label{sup_fig:bootstrap_distributions_for_simulated_sample_3}
  \end{center}
\end{figure}

@@

\begin{figure}[H]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/bootstrap_distributions_for_natural_sample_1_with_90_vpm_sliding.pdf}
  \caption{
  Bootstrap distributions of the mean distance to the future for natural populations.
  Empirical distances to the future were sampled with replacement and mean values for each bootstrap sample were calculated across n=10,000 bootstrap iterations.
  The horizontal gray line indicates the median of the bootstrap distribution for the naive model.
  Each model is annotated by the p-value from a bootstrap hypothesis test between the mean distances for that model and the naive model (see Methods).
  }
  \label{sup_fig:bootstrap_distributions_for_natural_sample_1_with_90_vpm_sliding}
  \end{center}
\end{figure}

@@

\begin{table}[H]
  \begin{center}
    \scalebox{0.9}{
        \input{tables/complete_natural_model_selection.tex}
    }
    \caption{
      All model coefficients and performance on validation and test data for natural populations ordered from best to worst by distance to the future, as in Table~\ref{table_simulated_model_selection}.
      Distances annotated with asterisks (*) were significantly closer to the future than the naive model as measured by bootstrap hypothesis tests (see Methods).
      Validation results are based on 23 timepoints.
      Test results are based on eight timepoints not observed during model training and validation.
      Model results for additional variants of fitness metrics including those based on epitope mutations and DMS preferences are included for reference.
    }
    \label{sup_table:complete_natural_model_selection}
  \end{center}
\end{table}
